---
title: "max_drivers_vs_time"
output: pdf_document
---


```{r}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message = FALSE, warning= FALSE}
library(ggplot2)
library(lubridate)
library(dplyr)
library(tidyverse)
library(stringr)
library(forcats)
library(hrbrthemes)
library(textclean)
library(xts)

setwd("/home/aquatic_max/neon4cast-aquatics/neon4cast-aquatics")

cleaned_data <- read.csv("~/neon4cast-aquatics/neon4cast-aquatics/data_exploration/cleaned_data.csv")

#colnames(cleaned_data)
 # "startDateTime"        "siteID"               "dissolvedOxygen"      "TFPrecipBulk"         "buoyWindSpeedMean"    "staPresMean"         
 # "tempRHMean"           "uPARMean"             "PARMean"              "surfacewaterElevMean" "groundwaterTempMean"  "surfWaterNitrateMean"


```


```{r, echo = FALSE, message = FALSE, warning= FALSE}
#Add month and year string columns

#  Convert to date if not already
cleaned_data$startDateTime <- as.Date(cleaned_data$startDateTime)

#  Get months
cleaned_data$Month <- substring(cleaned_data$startDateTime,6,7)

#  Get years
cleaned_data$Year <- substring(cleaned_data$startDateTime,1,4)

#Combine the Month and Year Column for graphing prep
cleaned_data$month_year <- as.yearmon(paste(cleaned_data$Month, cleaned_data$Year), "%m %Y")


#  Aggregate column of interest based on each month through time and get mean
monthly_avg_dissolvedOxygen <- aggregate(dissolvedOxygen ~ month_year, cleaned_data , mean )
monthly_avg_TFPrecipBulk <- aggregate(TFPrecipBulk ~ month_year, cleaned_data , mean )
monthly_avg_buoyWindSpeedMean <- aggregate(buoyWindSpeedMean ~ month_year, cleaned_data , mean )
monthly_avg_staPresMean <- aggregate(staPresMean ~ month_year, cleaned_data , mean )
monthly_avg_tempRHMean <- aggregate(tempRHMean ~ month_year, cleaned_data , mean )
monthly_avg_uPARMean <- aggregate(uPARMean ~ month_year, cleaned_data , mean )
monthly_avg_PARMean <- aggregate(PARMean ~ month_year, cleaned_data , mean )
monthly_avg_surfacewaterElevMean <- aggregate(surfacewaterElevMean ~ month_year, cleaned_data , mean )
monthly_avg_groundwaterTempMean <- aggregate(groundwaterTempMean ~ month_year, cleaned_data , mean )
monthly_avg_surfWaterNitrateMean <- aggregate(surfWaterNitrateMean ~ month_year, cleaned_data , mean )

# Multiyear monthly averages
yearly_avg_dissolvedOxygen <- aggregate(dissolvedOxygen ~ Month, cleaned_data , mean )
yearly_avg_TFPrecipBulk <- aggregate(TFPrecipBulk ~ Month, cleaned_data , mean )
yearly_avg_buoyWindSpeedMean <- aggregate(buoyWindSpeedMean ~ Month, cleaned_data , mean )
yearly_avg_staPresMean <- aggregate(staPresMean ~ Month, cleaned_data , mean )
yearly_avg_tempRHMean <- aggregate(tempRHMean ~ Month, cleaned_data , mean )
yearly_avg_uPARMean <- aggregate(uPARMean ~ Month, cleaned_data , mean )
yearly_avg_PARMean <- aggregate(PARMean ~ Month, cleaned_data , mean )
yearly_avg_surfacewaterElevMean <- aggregate(surfacewaterElevMean ~ Month, cleaned_data , mean )
yearly_avg_groundwaterTempMean <- aggregate(groundwaterTempMean ~ Month, cleaned_data , mean )
yearly_avg_surfWaterNitrateMean <- aggregate(surfWaterNitrateMean ~ Month, cleaned_data , mean )

```

## dissolvedOxygen
### Highest DO at start of year, decreases through November
```{r, echo = FALSE, message = FALSE, warning= FALSE}
#Average Mean Per Month
ggplot(monthly_avg_dissolvedOxygen, aes(x=month_year, y=dissolvedOxygen))+
  geom_point(aes(color = dissolvedOxygen),size = 3)+
  labs(title = "Average Dissolved Oxygen per Month",subtitle = "Sourced from NEON",y= "Dissolved Oxygen",x="Month & Year", color="DO Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))


#Multiyear Monthly Average
ggplot(yearly_avg_dissolvedOxygen, aes(x=fct_inorder(Month), y=dissolvedOxygen))+
  geom_point(aes(color = dissolvedOxygen),size = 3)+
  labs(title = "Multiyear Monthly Avg for Dissolved Oxygen",subtitle = "Sourced from NEON",y= "Dissolved Oxygen",x="Month", color="DO Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))

```
## TFPrecipBulk
```{r, echo = FALSE, message = FALSE, warning= FALSE}
#Average Mean Per Month
ggplot(monthly_avg_TFPrecipBulk, aes(x=month_year, y=TFPrecipBulk))+
  geom_point(aes(color = TFPrecipBulk),size = 3)+
  labs(title = "Average TFPrecipBulk per Month",subtitle = "Sourced from NEON",y= "TFPrecipBulk",x="Month & Year", color="TFPrecipBulk Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))


#Multiyear Monthly Average
ggplot(yearly_avg_TFPrecipBulk, aes(x=fct_inorder(Month), y=TFPrecipBulk))+
  geom_point(aes(color = TFPrecipBulk),size = 3)+
  labs(title = "Multiyear Monthly Avg for TFPrecipBulk",subtitle = "Sourced from NEON",y= "TFPrecipBulk",x="Month", color="TFPrecipBulk Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))

```

## buoyWindSpeedMean
```{r, echo = FALSE, message = FALSE, warning= FALSE}
#Average Mean Per Month
ggplot(monthly_avg_buoyWindSpeedMean, aes(x=month_year, y=buoyWindSpeedMean))+
  geom_point(aes(color = buoyWindSpeedMean),size = 3)+
  labs(title = "Average buoyWindSpeedMean per Month",subtitle = "Sourced from NEON",y= "buoyWindSpeedMean",x="Month & Year", color="buoyWindSpeedMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))


#Multiyear Monthly Average
ggplot(yearly_avg_buoyWindSpeedMean, aes(x=fct_inorder(Month), y=buoyWindSpeedMean))+
  geom_point(aes(color = buoyWindSpeedMean),size = 3)+
  labs(title = "Multiyear Monthly Avg for buoyWindSpeedMean",subtitle = "Sourced from NEON",y= "buoyWindSpeedMean",x="Month", color="buoyWindSpeedMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))

```

## staPresMean
### Strong Summer/Winter data separation
```{r, echo = FALSE, message = FALSE, warning= FALSE}
#Average Mean Per Month
ggplot(monthly_avg_staPresMean, aes(x=month_year, y=staPresMean))+
  geom_point(aes(color = staPresMean),size = 3)+
  labs(title = "Average staPresMean per Month",subtitle = "Sourced from NEON",y= "staPresMean",x="Month & Year", color="staPresMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))


#Multiyear Monthly Average
ggplot(yearly_avg_staPresMean, aes(x=fct_inorder(Month), y=staPresMean))+
  geom_point(aes(color = staPresMean),size = 3)+
  labs(title = "Multiyear Monthly Avg for staPresMean",subtitle = "Sourced from NEON",y= "staPresMean",x="Month", color="staPresMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))

```

## tempRHMean
### Summer increase
```{r, echo = FALSE, message = FALSE, warning= FALSE}
#Average Mean Per Month
ggplot(monthly_avg_tempRHMean, aes(x=month_year, y=tempRHMean))+
  geom_point(aes(color = tempRHMean),size = 3)+
  labs(title = "Average tempRHMean per Month",subtitle = "Sourced from NEON",y= "tempRHMean",x="Month & Year", color="tempRHMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))


#Multiyear Monthly Average
ggplot(yearly_avg_tempRHMean, aes(x=fct_inorder(Month), y=tempRHMean))+
  geom_point(aes(color = tempRHMean),size = 3)+
  labs(title = "Multiyear Monthly Avg for tempRHMean",subtitle = "Sourced from NEON",y= "tempRHMean",x="Month", color="tempRHMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))

```

## uPARMean
```{r, echo = FALSE, message = FALSE, warning= FALSE}
#Average Mean Per Month
ggplot(monthly_avg_uPARMean, aes(x=month_year, y=uPARMean))+
  geom_point(aes(color = uPARMean),size = 3)+
  labs(title = "Average uPARMean per Month",subtitle = "Sourced from NEON",y= "uPARMean",x="Month & Year", color="uPARMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))


#Multiyear Monthly Average
ggplot(yearly_avg_uPARMean, aes(x=fct_inorder(Month), y=uPARMean))+
  geom_point(aes(color = uPARMean),size = 3)+
  labs(title = "Multiyear Monthly Avg for uPARMean",subtitle = "Sourced from NEON",y= "uPARMean",x="Month", color="uPARMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))

```

## PARMean
```{r, echo = FALSE, message = FALSE, warning= FALSE}
#Average Mean Per Month
ggplot(monthly_avg_PARMean, aes(x=month_year, y=PARMean))+
  geom_point(aes(color = PARMean),size = 3)+
  labs(title = "Average PARMean per Month",subtitle = "Sourced from NEON",y= "PARMean",x="Month & Year", color="PARMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))


#Multiyear Monthly Average
ggplot(yearly_avg_PARMean, aes(x=fct_inorder(Month), y=PARMean))+
  geom_point(aes(color = PARMean),size = 3)+
  labs(title = "Multiyear Monthly Avg for PARMean",subtitle = "Sourced from NEON",y= "PARMean",x="Month", color="PARMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))

```

## surfacewaterElevMean
### Decrease in surface water elev through time
```{r, echo = FALSE, message = FALSE, warning= FALSE}
#Average Mean Per Month
ggplot(monthly_avg_surfacewaterElevMean, aes(x=month_year, y=surfacewaterElevMean))+
  geom_point(aes(color = surfacewaterElevMean),size = 3)+
  labs(title = "Average surfacewaterElevMean per Month",subtitle = "Sourced from NEON",y= "surfacewaterElevMean",x="Month & Year", color="surfacewaterElevMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))


#Multiyear Monthly Average
ggplot(yearly_avg_surfacewaterElevMean, aes(x=fct_inorder(Month), y=surfacewaterElevMean))+
  geom_point(aes(color = surfacewaterElevMean),size = 3)+
  labs(title = "Multiyear Monthly Avg for surfacewaterElevMean",subtitle = "Sourced from NEON",y= "surfacewaterElevMean",x="Month", color="surfacewaterElevMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))

```

## groundwaterTempMean
### Steady increase in groundwater temp through time
```{r, echo = FALSE, message = FALSE, warning= FALSE}
#Average Mean Per Month
ggplot(monthly_avg_groundwaterTempMean, aes(x=month_year, y=groundwaterTempMean))+
  geom_point(aes(color = groundwaterTempMean),size = 3)+
  labs(title = "Average groundwaterTempMean per Month",subtitle = "Sourced from NEON",y= "groundwaterTempMean",x="Month & Year", color="groundwaterTempMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))


#Multiyear Monthly Average
ggplot(yearly_avg_groundwaterTempMean, aes(x=fct_inorder(Month), y=groundwaterTempMean))+
  geom_point(aes(color = groundwaterTempMean),size = 3)+
  labs(title = "Multiyear Monthly Avg for groundwaterTempMean",subtitle = "Sourced from NEON",y= "groundwaterTempMean",x="Month", color="groundwaterTempMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))

```

## surfWaterNitrateMean
### Slight increase in nitrate through time
```{r, echo = FALSE, message = FALSE, warning= FALSE}
#Average Mean Per Month
ggplot(monthly_avg_surfWaterNitrateMean, aes(x=month_year, y=surfWaterNitrateMean))+
  geom_point(aes(color = surfWaterNitrateMean),size = 3)+
  labs(title = "Average surfWaterNitrateMean per Month",subtitle = "Sourced from NEON",y= "surfWaterNitrateMean",x="Month & Year", color="surfWaterNitrateMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))


#Multiyear Monthly Average
ggplot(yearly_avg_surfWaterNitrateMean, aes(x=fct_inorder(Month), y=surfWaterNitrateMean))+
  geom_point(aes(color = surfWaterNitrateMean),size = 3)+
  labs(title = "Multiyear Monthly Avg for surfWaterNitrateMean",subtitle = "Sourced from NEON",y= "surfWaterNitrateMean",x="Month", color="surfWaterNitrateMean Level")+
  theme(axis.text.x = element_text(angle = 40, vjust = 1, hjust=1))

```