---
title: "exploration"
output: html_document
---

# In this first Part, I import all the relevant data from NEON;  data exploration stuff starts at line 241 
```{r}
library(neonstore)
library(dplyr)
library(tidyverse)
library(lubridate)
library(neonUtilities)

Sys.setenv("NEONSTORE_HOME" = "neon_store/")
if(dir.exists("neon_store/")){
  dir.create("neon_store/")
}
```

Pulling all the data that I think will be relevant from NEON
```{r}
focal_sites <- c("BARC", "POSE")
oxy <- neonUtilities::loadByProduct("DP1.20288.001", site = focal_sites) # Water quality (includes DO)
precipitation <- neonUtilities::loadByProduct("DP1.00006.001", site =  focal_sites) # precipitation
wind <- neonUtilities::loadByProduct("DP1.20059.001", site =  focal_sites) # wind speed
pressure <- neonUtilities::loadByProduct("DP1.20004.001", site =  focal_sites) # baromeVtric pressure
temperature <- neonUtilities::loadByProduct("DP1.20046.001", site =  focal_sites) # temperature
par_surface <- neonUtilities::loadByProduct("DP1.20042.001", site =  focal_sites) # PAR at surface
par_below <- neonUtilities::loadByProduct("DP1.20261.001", site =  focal_sites) # PAR below surface
elevation_surface <- neonUtilities::loadByProduct("DP1.20016.001", site =  focal_sites) # Elevation of surf.
groundwater_temp <- neonUtilities::loadByProduct("DP1.20217.001", site =  focal_sites) # Groundwater Temp
nitrate_surface <- neonUtilities::loadByProduct("DP1.20033.001", site =  focal_sites) # Nitrate at surf.
```
Tidying up DO data, for exploration purposes I'm only going to look at the means here on but for forecasting, dealing with uncertainty will probably be the way to go.
```{r}
oxy <- oxy$waq_instantaneous %>%
  filter(is.na(dissolvedOxygen) != TRUE)

oxy_cleaned <- oxy %>%
  dplyr::select(siteID, startDateTime, sensorDepth, dissolvedOxygen,
                dissolvedOxygenExpUncert,dissolvedOxygenFinalQF) %>%
  dplyr::filter(dissolvedOxygenFinalQF == 0,
                sensorDepth > 0) %>%
  dplyr::mutate(startDateTime = as_datetime(startDateTime)) %>%
  dplyr::mutate(date = as_date(startDateTime),
                hour = hour(startDateTime)) %>%
  dplyr::group_by(siteID, date, hour) %>%
  dplyr::summarize(sensorDepth = mean(sensorDepth, na.rm = TRUE),
                   dissolvedOxygen = mean(dissolvedOxygen, na.rm = TRUE),
                   dissolvedOxygenExpUncert = mean(dissolvedOxygenExpUncert, na.rm = TRUE),
                   sensorDepth = mean(sensorDepth, na.rm = TRUE), .groups = "drop") %>%
  dplyr::mutate(startDateTime = make_datetime(year = year(date), month = month(date),
                                              day = day(date), hour = hour,
                                              min = 0, tz ="UTC")) %>%
  dplyr::select(siteID, startDateTime, sensorDepth, dissolvedOxygen, dissolvedOxygenExpUncert)
```

Cleaning up precipitation data and merging with DO table.
```{r}
site_change <- function(x) {
  if (x == "OSBS"){return("BARC")}
  else{return("POSE")}
}

precipitation_cleaned <- precipitation$THRPRE_30min %>%
  dplyr::select(siteID, startDateTime, TFPrecipBulk) %>%
  dplyr::mutate(startDateTime = as_datetime(startDateTime)) %>%
  dplyr::mutate(date = as_date(startDateTime),
                hour = hour(startDateTime)) %>%
  dplyr::group_by(siteID, date, hour) %>%
  dplyr::summarize(TFPrecipBulk = mean(TFPrecipBulk, na.rm = TRUE)) %>%
  dplyr::mutate(startDateTime = make_datetime(year = year(date), month = month(date),
                                              day = day(date), hour = hour,
                                              min = 0, tz ="UTC")) %>%
  dplyr::select(siteID, startDateTime, TFPrecipBulk) %>%
  dplyr::mutate(siteID = site_change(siteID))

sensorData <- merge(oxy_cleaned, precipitation_cleaned, by=c("startDateTime", "siteID"))
```
Cleaning up and merging wind data
```{r}
wind_cleaned <- wind$WSDBuoy_30min %>%
  select(siteID, startDateTime, buoyWindSpeedMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(buoyWindSpeedMean))) %>%
  summarise(buoyWindSpeedMean = mean(buoyWindSpeedMean, na.rm = TRUE))  %>%
  mutate(startDateTime = make_datetime(year = year(date), month = month(date),
                                              day = day(date), hour = hour,
                                              min = 0, tz ="UTC")) %>%
  select(siteID, startDateTime, buoyWindSpeedMean)
  
sensorData <- sensorData %>% 
  merge(wind_cleaned, by=c("startDateTime", "siteID"))
```

Cleaning up and merging pressure data
```{r}
pressure_cleaned <- pressure$BP_30min %>%
  select(siteID, startDateTime, staPresMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(staPresMean))) %>%
  summarise(staPresMean = mean(staPresMean, na.rm = TRUE))  %>%
  mutate(startDateTime = make_datetime(year = year(date), month = month(date),
                                              day = day(date), hour = hour,
                                              min = 0, tz ="UTC")) %>%
  select(siteID, startDateTime, staPresMean)
  
sensorData <- sensorData %>% 
  merge(pressure_cleaned, by=c("startDateTime", "siteID"))
```

Cleaning up and merging temperature data
```{r}
temperature_cleaned <- temperature$RHbuoy_30min %>%
  select(siteID, startDateTime, tempRHMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(tempRHMean))) %>%
  summarise(tempRHMean = mean(tempRHMean, na.rm = TRUE))  %>%
  mutate(startDateTime = make_datetime(year = year(date), month = month(date),
                                              day = day(date), hour = hour,
                                              min = 0, tz ="UTC")) %>%
  select(siteID, startDateTime, tempRHMean)
  
sensorData <- sensorData %>% 
  merge(temperature_cleaned, by=c("startDateTime", "siteID"))
```

Cleaning up and merging surface water radiation data thats relevant to photosynthesis
```{r}
par_surface_cleaned <- par_surface$PARWS_30min %>%
  select(siteID, startDateTime, PARMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(PARMean))) %>%
  summarise(PARMean = mean(PARMean, na.rm = TRUE))  %>%
  mutate(startDateTime = make_datetime(year = year(date), month = month(date),
                                              day = day(date), hour = hour,
                                              min = 0, tz ="UTC")) %>%
  select(siteID, startDateTime, PARMean)
  
sensorData <- sensorData %>% 
  merge(par_surface_cleaned, by=c("startDateTime", "siteID"))
```

Cleaning up and merging below water radiation data thats relevant to photosynthesis
```{r}
par_below_cleaned <- par_below$uPAR_30min %>%
  select(siteID, startDateTime, uPARMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(uPARMean))) %>%
  summarise(uPARMean = mean(uPARMean, na.rm = TRUE))  %>%
  mutate(startDateTime = make_datetime(year = year(date), month = month(date),
                                              day = day(date), hour = hour,
                                              min = 0, tz ="UTC")) %>%
  select(siteID, startDateTime, uPARMean)
  
sensorData <- sensorData %>% 
  merge(par_below_cleaned, by=c("startDateTime", "siteID"))
```

Cleaning up and merging elevation data
```{r}
elevation_surface_cleaned <- elevation_surface$EOS_30_min %>%
  select(siteID, startDateTime, surfacewaterElevMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(surfacewaterElevMean))) %>%
  summarise(surfacewaterElevMean = mean(surfacewaterElevMean, na.rm = TRUE))  %>%
  mutate(startDateTime = make_datetime(year = year(date), month = month(date),
                                              day = day(date), hour = hour,
                                              min = 0, tz ="UTC")) %>%
  select(siteID, startDateTime, surfacewaterElevMean)
  
sensorData <- sensorData %>% 
  merge(elevation_surface_cleaned, by=c("startDateTime", "siteID"))
```

Cleaning up and merging groundwater temp data
```{r}
groundwater_temp_cleaned <- groundwater_temp$TGW_30_minute %>%
  select(siteID, startDateTime, groundwaterTempMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(groundwaterTempMean))) %>%
  summarise(groundwaterTempMean = mean(groundwaterTempMean, na.rm = TRUE))  %>%
  mutate(startDateTime = make_datetime(year = year(date), month = month(date),
                                              day = day(date), hour = hour,
                                              min = 0, tz ="UTC")) %>%
  select(siteID, startDateTime, groundwaterTempMean)
  
sensorData <- sensorData %>% 
  merge(groundwater_temp_cleaned, by=c("startDateTime", "siteID"))
```

```{r}
nitrate_surface_cleaned <- nitrate_surface$NSW_15_minute %>%
  select(siteID, startDateTime, surfWaterNitrateMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(surfWaterNitrateMean))) %>%
  summarise(surfWaterNitrateMean = mean(surfWaterNitrateMean, na.rm = TRUE))  %>%
  mutate(startDateTime = make_datetime(year = year(date), month = month(date),
                                              day = day(date), hour = hour,
                                              min = 0, tz ="UTC")) %>%
  select(siteID, startDateTime, surfWaterNitrateMean)
  
sensorData <- sensorData %>% 
  merge(nitrate_surface_cleaned, by=c("startDateTime", "siteID"))
```


```{r}
cleanedData <- sensorData %>%
  select(startDateTime, siteID, dissolvedOxygen, TFPrecipBulk, buoyWindSpeedMean, 
         staPresMean, tempRHMean, uPARMean, PARMean, surfacewaterElevMean,
         groundwaterTempMean, surfWaterNitrateMean)

write.csv(cleanedData, file='cleaned_data.csv', row.names = FALSE)
```

# Importing CSV and plotting!
Loading `cleanedData` csv
```{r}
impCleanedData <- read.csv("cleaned_data.csv")
```

Plotting variables against DO
```{r}
x_names <- names(impCleanedData)[4:ncol(impCleanedData)] 
for (i in x_names){
  plt <- impCleanedData %>% 
    ggplot(aes_string(x = i, y = "dissolvedOxygen")) +
    geom_point()
  print(plt)
}
```