---
title: "exploration"
output: html_document
---

# In this first Part, I import all the relevant data from NEON;  data exploration stuff starts at line 241 
```{r}
library(neonstore)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(lubridate)

```

```{r}

# Use a remote connection, so `dplyr` operations can be run by SQL:
db <- neon_db()
waq <- tbl(db, "waq_instantaneous-basic-DP1.20288.001")
# precip <- 
wind <- tbl(db, "WSDBuoy_30min-basic-DP1.20059.001")
pressure <- tbl(db, "BP_30min-basic-DP1.20004.001")
airT <- tbl(db, "RHbuoy_30min-basic-DP1.20046.001")
PARsurf <- tbl(db, "PARWS_30min-basic-DP1.20042.001")
PARbelow <- tbl(db, "uPAR_30min-basic-DP1.20261.001")
elevSurf <- tbl(db, "EOS_30_min-basic-DP1.20016.001")
waterT <- tbl(db, "TGW_30_minute-basic-DP1.20217.001")
nitrate <- tbl(db, "NSW_15_minute-basic-DP1.20033.001")
```

The rest is not going to work now! TO BE UPDATED! 

Tidying up DO data, for exploration purposes I'm only going to look at the means here on but for forecasting, dealing with uncertainty will probably be the way to go. Note that the water quality data has potentially more important drivers like chlorophyll, turbidity and pH.
```{r}
oxy_cleaned <- waq %>%
  select(siteID, startDateTime, dissolvedOxygen) %>%
  filter(!(is.na(dissolvedOxygen))) %>% 
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
                hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  summarise(dissolvedOxygen = mean(dissolvedOxygen, na.rm = TRUE)) 
```

Cleaning up precipitation data and merging with DO table. There is no precipitation data for the 2 sites -- NEONUtilities grabbed from proximal sites.
```{r}
# site_change <- function(x) {
#   if (x == "OSBS"){return("BARC")}
#   else{return("POSE")}
# }
# 
# precipitation_cleaned <- precipitation$THRPRE_30min %>%
#   dplyr::select(siteID, startDateTime, TFPrecipBulk) %>%
#   dplyr::mutate(startDateTime = as_datetime(startDateTime)) %>%
#   dplyr::mutate(date = as_date(startDateTime),
#                 hour = hour(startDateTime)) %>%
#   dplyr::group_by(siteID, date, hour) %>%
#   dplyr::summarize(TFPrecipBulk = mean(TFPrecipBulk, na.rm = TRUE)) %>%
#   dplyr::mutate(startDateTime = make_datetime(year = year(date), month = month(date),
#                                               day = day(date), hour = hour,
#                                               min = 0, tz ="UTC")) %>%
#   dplyr::select(siteID, startDateTime, TFPrecipBulk) %>%
#   dplyr::mutate(siteID = site_change(siteID))
# 
# sensorData <- inner_join(oxy_cleaned, precipitation_cleaned, by=c("startDateTime", "siteID"))
```

Cleaning up and merging wind data
```{r}
wind_cleaned <- wind %>% 
  select(siteID, startDateTime, buoyWindSpeedMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(buoyWindSpeedMean))) %>%
  summarise(buoyWindSpeedMean = mean(buoyWindSpeedMean, na.rm = TRUE))  %>%
  select(siteID, date, hour, buoyWindSpeedMean)
  
sensorData <- inner_join(oxy_cleaned, wind_cleaned, by=c("date", "hour", "siteID"))
```

Cleaning up and merging pressure data
```{r}
pressure_cleaned <- pressure %>%
  select(siteID, startDateTime, staPresMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(staPresMean))) %>%
  summarise(staPresMean = mean(staPresMean, na.rm = TRUE))  %>%
  select(siteID, date, hour, staPresMean)
  
sensorData <- sensorData %>% 
  inner_join(pressure_cleaned, by=c("date", "hour", "siteID"))
```

Cleaning up and merging temperature data
```{r}
airT_cleaned <- airT %>%
  select(siteID, startDateTime, tempRHMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(tempRHMean))) %>%
  summarise(tempRHMean = mean(tempRHMean, na.rm = TRUE))  %>%
  select(siteID, date, hour, tempRHMean)
  
sensorData <- sensorData %>% 
  inner_join(airT_cleaned, by=c("date", "hour", "siteID"))
```

Cleaning up and merging surface water radiation data thats relevant to photosynthesis
```{r}
PARsurf_cleaned <- PARsurf %>%
  select(siteID, startDateTime, PARMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(PARMean))) %>%
  summarise(PARMean = mean(PARMean, na.rm = TRUE))  %>%
  select(siteID, date, hour, PARMean)
  
sensorData <- sensorData %>% 
  inner_join(PARsurf_cleaned, by=c("date", "hour", "siteID"))
```

Cleaning up and merging below water radiation data thats relevant to photosynthesis
```{r}
PARbelow_cleaned <- PARbelow %>%
  select(siteID, startDateTime, uPARMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(uPARMean))) %>%
  summarise(uPARMean = mean(uPARMean, na.rm = TRUE))
  
sensorData <- sensorData %>% 
  inner_join(PARbelow_cleaned, by=c("date", "hour", "siteID"))
```

Cleaning up and merging elevation data
```{r}
elevSurf_cleaned <- elevSurf %>%
  select(siteID, startDateTime, surfacewaterElevMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(surfacewaterElevMean))) %>%
  summarise(surfacewaterElevMean = mean(surfacewaterElevMean, na.rm = TRUE))  %>%
  select(siteID, date, hour, surfacewaterElevMean)
  
sensorData <- sensorData %>% 
  inner_join(elevSurf_cleaned, by=c("date", "hour", "siteID"))
```

Cleaning up and merging groundwater temp data
```{r}
waterT_cleaned <- waterT %>%
  select(siteID, startDateTime, groundwaterTempMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(groundwaterTempMean))) %>%
  summarise(groundwaterTempMean = mean(groundwaterTempMean, na.rm = TRUE))  %>%
  select(siteID, date, hour, groundwaterTempMean)
  
sensorData <- sensorData %>% 
  inner_join(waterT_cleaned, by=c("date", "hour", "siteID"))
```

```{r}
nitrate_cleaned <- nitrate %>%
  select(siteID, startDateTime, surfWaterNitrateMean) %>%
  mutate(startDateTime = as_datetime(startDateTime)) %>%
  mutate(date = as_date(startDateTime),
         hour = hour(startDateTime)) %>%
  group_by(siteID, date, hour) %>%
  filter(!(is.na(surfWaterNitrateMean))) %>%
  summarise(surfWaterNitrateMean = mean(surfWaterNitrateMean, na.rm = TRUE))  %>%
  select(siteID, date, hour, surfWaterNitrateMean)
  
sensorData <- sensorData %>% 
  inner_join(nitrate_cleaned, by=c("date", "hour", "siteID"))
```


```{r}
cleanedData <- sensorData %>%
  select(date, hour, siteID, dissolvedOxygen, buoyWindSpeedMean, 
         staPresMean, tempRHMean, uPARMean, PARMean, surfacewaterElevMean,
         groundwaterTempMean, surfWaterNitrateMean)

#write.csv(cleanedData, file='cleaned_data.csv', row.names = FALSE)
```

# Importing CSV and plotting!
Loading `cleanedData` csv
```{r}
#impCleanedData <- read.csv("cleaned_data.csv")
```

Plotting variables against DO
```{r}
plot <- cleanedData %>% collect() %>%
  pivot_longer(colnames(cleanedData)[5:ncol(cleanedData)], names_to = "Driver", values_to = "Obs") %>% 
  ggplot() +
  geom_point(mapping = aes(x = Obs, y = dissolvedOxygen)) +
  facet_wrap(~ Driver, nrow = 3, scales = "free")
plot
```